### 2018-04-23 contents/abstract.tex
Updates from ShareLaTeX

\chapter{Abstract}

Idiopathic (Classic) Trigeminal Neuralgia (TN) is a facial neuropathic pain syndrome characterized by paroxysmal, shock-like pain condition affecting one or more of the three trigeminal nerve (CNV) branches. TN is believed to be associated with nerve-vascular compression in the CNV root-entry-zone, but its pathophysiology is still unclear. Single-tensor diffusion tensor neuroimaging (DTI) studies of CN V revealed diffusivity changes in the cistern segment. However, the portion of the nerve within the brainstem remained elusive due to the limits of DTI, and the lack of reliable software to researchers. Thereby the study of TN using diffusion imaging is time-consuming and error-prone due to the complete manual involvement in all steps of data process, placement of region-of-interest (ROI), and tractography parameter tuning, and data analysis. 
There are two aims of this thesis: 1) to develop a fully-automated software framework to collate and analyze diffusion tractography to reduce human error and increase iteration speed of experiments. 2) to apply the method developed to the analysis of the trigeminal system in a more substantial patient group, and apply state-of-the-art diffusion imaging methods to advance the study of TN further. 
The specific aims are: a) establish the feasibility of high angular resolution diffusion tractography of brainstem CNV; b) establish the best way to minimize DWI to T1 co-registration error allow non-linear registration of tractography across multiple subjects; c) Create the software framework to generate and quantify tractography at the group level; d) apply the methodology to the reconstruction and quantification of the trigeminal sensory pathway. 
Towards these goals, in Study I, we establish the feasibility of applying spherical deconvolution tractography to delineate the full course of CNV from the trigeminal nucleus and demonstrate that TN is uniquely identified by disruptions in the cistern/REZ, while MS-TN can be distinguished by disruptions in the brainstem course of the nerve. In Study II, we determine that the best T1-DWI co-registration scalar is the Mean DWI image, and that the registration can potentially be improved with a novel spherical harmonic parameter scalar imaged called Anisotropic Power image. In Study III, we present the Selective Automated Group Integrated Tractography (SAGIT) software framework. Finally, in Study IV, we deploy end-to-end machine-learning TN classification to auto-discovers diffusivity disruptions in the cistern/REZ CNV, the trigeminopontothalamic decussaion, and thalamocortico S1 pathway. 
In sum, this thesis presents a detailed road-map of both the development and application of a novel end-to-end diffusion tractography machine learning classification. The application of this system to TN revealed specific diffusivity changes in all three segments of the trigeminal CN V, pontine decussation, and S1  white matter pathways, and pin-points the exact location of the diffusivity disruption at millimetre level.

>>>>

\chapter{Abstract}
% \begin{abstract}
\doublespacing
Idiopathic (Classic) Trigeminal Neuralgia (TN) is a facial neuropathic pain syndrome characterized by paroxysmal, shock-like pain condition affecting one or more of the three trigeminal nerve (CNV) branches. TN is believed to be associated with nerve-vascular compression in the CNV root-entry-zone, but its pathophysiology is still unclear. Single-tensor diffusion tensor neuroimaging (DTI) studies of CN V revealed diffusivity changes in the cistern segment. However, the portion of the nerve within the brainstem remained elusive due to the limits of DTI, and the lack of reliable software to researchers. Thereby the study of TN using diffusion imaging is time-consuming and error-prone due to the complete manual involvement in all steps of data process, placement of region-of-interest (ROI), and tractography parameter tuning, and data analysis. 
There are two aims of this thesis: 1) to develop a fully-automated software framework to collate and analyze diffusion tractography to reduce human error and increase iteration speed of experiments. 2) to apply the method developed to the analysis of the trigeminal system in a more substantial patient group, and apply state-of-the-art diffusion imaging methods to advance the study of TN further. 
The specific aims are: a) establish the feasibility of high angular resolution diffusion tractography of brainstem CNV; b) establish the best way to minimize DWI to T1 co-registration error allow non-linear registration of tractography across multiple subjects; c) Create the software framework to generate and quantify tractography at the group level; d) apply the methodology to the reconstruction and quantification of the trigeminal sensory pathway. 
Towards these goals, in Study I, we establish the feasibility of applying spherical deconvolution tractography to delineate the full course of CNV from the trigeminal nucleus and demonstrate that TN is uniquely identified by disruptions in the cistern/REZ, while MS-TN can be distinguished by disruptions in the brainstem course of the nerve. In Study II, we determine that the best T1-DWI co-registration scalar is the Mean DWI image, and that the registration can potentially be improved with a novel spherical harmonic parameter scalar imaged called Anisotropic Power image. In Study III, we present the Selective Automated Group Integrated Tractography (SAGIT) software framework. Finally, in Study IV, we deploy end-to-end machine-learning TN classification to auto-discovers diffusivity disruptions in the cistern/REZ CNV, the trigeminopontothalamic decussaion, and thalamocortico S1 pathway. 
In sum, this thesis presents a detailed road-map of both the development and application of a novel end-to-end diffusion tractography machine learning classification. The application of this system to TN revealed specific diffusivity changes in all three segments of the trigeminal CN V, pontine decussation, and S1  white matter pathways, and pin-points the exact location of the diffusivity disruption at millimetre level.
% \end{abstract}

### 2018-04-23 contents/hypotheses.tex
Updates from ShareLaTeX

\subsection{Specific Hypothesis}
\begin{itemize}
    \item TN is known to relate to vascular compression of the cistern REZ, therefore we expect its diffusivity signature to occur in the cisternal region.
    \item MS-TN should be related to brainstem MS lesions. Therefore CN V segments close to the lesions are expected to show more significant changes in tissue diffusivity, whereas segments far away from the lesions will not. 
    \item As a demyelinating disease, we expect to find a pattern of change in diffusivity signatures that reflects demyelination, including increase in RD, and decrease in FA. 
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item Symmetric-diffeomorphic registration will result in the best accuracy, but with limited asymptotic improvement over affine registration alone.
    \item Averaged DWI should be the best scalar image, as it provides a balanced contrast between white and grey matter.
    \item Anisotropic-power (AP) image will show more substantial improvement over the existing scalar images.
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item DTI is known to not able to delineate cross-fibers regions. If the group tractography works consistently, then its output should reflect such known algorithmic limitations, and DTI group tracts should not be able to perform well on more complex anatomies. The grouped tracts should however show consistent regional anatomical overlap across individuals.
    \item By merging the tracts of many individuals, the tractography delineations can complement each other, and this may improve the anatomical fidelity over individual cases.
    \item Specific tractography algorithms may be preferentially suited for the visualization of specific areas in the brain. For examples, deterministic tractography should preferentially be more suitable for nerve tractography, and probabilistic method better for deep neural pathways such as the optic and auditory radiations.
    \item NOS score should correlate with expert ratings, and show improved rating consistency over human rater variability.
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item To validate the study, CN V diffusivity signature should match findings from Study I.
    \item At the level of CN V, the affected cistern segment should show differentiation from the unaffected side.
    \item The affected S1 segment should differ from the unaffected side.
    \item The affected TPT segment might show more differentiation near the end points, due to possible differences in trigeminal nucleus and thalamus diffusivity. However, its mid-segment diffusivity might be inconclusive, as the diffusivity metrics such as FA and RD might not be adequate in describing the complex crossings at the pontine decussation. 
\end{itemize}

>>>>

\subsection{Specific Hypothesis}
\begin{itemize}
    \item TN is known to relate to vascular compression of the cistern REZ, therefore its diffusivity signature will occur in the cisternal region.
    \item MS-TN is related to brainstem MS lesions. Therefore CN V segments close to the lesions will show more significant changes in tissue diffusivity, whereas segments far away from the lesions will not. 
    \item As a demyelinating disease, the pattern of change for MS-TN diffusivity signatures will reflects demyelination, including increase in RD, and decrease in FA. 
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item Symmetric-diffeomorphic registration will result in the best accuracy, but with limited asymptotic improvement over affine registration alone.
    \item Averaged DWI will provide good registration quality, as it provides a balanced contrast between white and grey matter.
    \item Anisotropic-power (AP) image will show more substantial improvement over the existing scalar images.
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item DTI is known to not able to delineate cross-fibers regions. If the group tractography works consistently, then its output will reflect such known algorithmic limitations, and DTI group tracts will not be able to perform well on more complex anatomies. The grouped tracts will however show consistent regional anatomical overlap across individuals.
    \item Specific tractography algorithms will be preferentially suited for the visualization of specific areas in the brain. For examples, deterministic tractography will preferentially be more suitable for nerve tractography, and probabilistic method better for deep neural pathways such as the optic and auditory radiations.
    \item NOS score will correlate with expert ratings, and show improved rating consistency over human rater variability.
\end{itemize}

@@

\subsection{Specific Hypothesis}
\begin{itemize}
    \item To validate the study, CN V diffusivity signature will match findings from Study I.
    \item At the level of CN V, the affected cistern segment will show differentiation from the unaffected side.
    \item The affected S1 segment will differ from the unaffected side.
    \item The affected TPT segment will show more differentiation near the end points, due to possible differences in trigeminal nucleus and thalamus diffusivity. However, its mid-segment diffusivity willht be inconclusive, as the diffusivity metrics such as FA and RD might not be adequate in describing the complex crossings at the pontine decussation. 
\end{itemize}
